Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex
Switch from conjunction to comma in tensor product definition
added examples to eval

\textbf{Single action} specifications consist of no more than one 
forward, backward, or centered region and can be combined with any
number of pointed regions. We identified 4,837 single-action
specifications, of which 1,532 were single-action with a
\texttt{nonpointed} modifier.

@@

>>>>

\textbf{Single-action} specifications consist of no more than one 
forward, backward, or centered region and can be combined with any
number of pointed regions. We identified 4,837 single-action
specifications, of which 1,532 were single-action with a
\texttt{nonpointed} modifier.

@@

\begin{figure}[t]\begin{minted}[fontsize=\scriptsize,breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,nonpointed))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(pointed(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2,nonpointed))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(pointed(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,nonpointed))*(pointed(dim=3))+(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2))*(pointed(dim=3))::x
\end{minted}
\caption{A complex specification inferred from
  \textbf{Unified Model}\label{fig:smagorinsky}}
\vspace{-1em}
\end{figure}

The single- and multi-action classes represent more complex stencils
with a real possibility for programmer error. For example, the Unified
Model has an implementation of the Smagorinsky subgrid-scale model for
calculating turbulence on which our inference yields 39 specs from 340
lines of code. This is an impressive reduction given
the complexity of the algorithm.  We show one example in
Figure~\ref{fig:smagorinsky}. This specifies the access pattern to
a 3-dimensional array (which we have renamed to be called
\texttt{x}). The array is accessed 8 times in the computation.

\subsection{Detecting errors in the 2-D Jacobi iteration}

One common example of a stencil computation is two-dimensional
Jacobi iteration that repeatedly goes through each cell in a matrix
and computes the average value of the four adjacent cells. The kernel
can be expressed in a line of Fortran code:
\begin{minted}{fortran}
  a(i,j) = (a(i-1,j)+a(i+1,j)+a(i,j+1)+a(i,j-1))/4
\end{minted}
and CamFort is able to infer a precise specification for this:
\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  != stencil pointed(dim=1)*centered(depth=1,dim=2,nonpointed)+pointed(dim=2)*centered(depth=1,dim=1,nonpointed)::a
\end{minted}

We examined whether programmer errors would be detected by replacing
the array index offsets with $-1$, $0$, or $1$ and running our
verification algorithm. Camfort correctly reported a verification
failure in each of 6,537 permutations corresponding to an error.  We
note that the iteration is computing the average of four adjacent
cells so $24$ ($4!$) of the possible array index pertubations do not
correspond to a programming error.

% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught
